---
title: "Introduction to GPcov"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to GPcov}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(GPcov)
```

This is a quick introduction of how to use the `GPCov` package. 

## `compute_cov_matrix()` function

In `GPcov`, the main function we will use is named `compute_cov_matrix()`, and it has two arguments, namely `points` and `cov_func`.

`points` has to be a `data.frame` with `n` rows and `d` columns, such that it corresponds to `n` `d`-dimensional points. For instance, we can define it a set of $30 \times 30 = 900$ points in $\mathbb{R}^2$.
```{r}
x <- seq(from = 1, to = 30, by = 1)
y <- seq(from = 1, to = 30, by = 1)
z <- expand.grid(x, y)
print(class(z))
print(nrow(z))
```

On the other hand, `cov_func` has a few options currently implemented, namely `white_noise`, `brownian_motion`, `exponential`, and `matern_model`. Each of these models correspond to a specific covariance function, which we will detail now.

Let $X(\mathbf{u})$ be a random process, such that $\mathbf{u} \in \mathbb{R}^d$, for some positive integer $d$. Also, whenever it is used, let $h = ||\mathbf{u}_1 - \mathbf{u}_2||$ denote the Euclidean distance between $\mathbf{u}_1$ and $\mathbf{u}_2$, where $\mathbf{u}_1, \mathbf{u}_2 \in \mathbb{R}^d$. Finally, let $\text{Cov}(X(\mathbf{u}_1), X(\mathbf{u}_2))$ denote the covariance between $X(\mathbf{u}_1)$ and $\mathbf{u}_2$. That being defined, the available models are specified in the following way

- `white_noise`:
$$
\text{Cov}(X(\mathbf{u}_1), X(\mathbf{u}_2)) = \left\{
\begin{array}{ll}
0 & \text{ if } \mathbf{u}_1 \neq \mathbf{u}_2 \\
1 & \text{ if } \mathbf{u}_1 = \mathbf{u}_2
\end{array}
\right.
$$
- `brownian_motion`: It is only defined for $u \in \mathbb{R}$; in that case,
$$
\text{Cov}(X(u_1), X(u_2)) = \min(u_1, u_2)
$$

- `exponential`: for $\sigma^2, \beta > 0$,
$$
\text{Cov}(X(u_1), X(u_2)) = \text{Cov}(h) = \sigma^2 \cdot \exp\{-h/\beta\}
$$
- `matern_model`: for $\sigma^2, \nu, \beta > 0$,
$$
\text{Cov}(X(u_1), X(u_2)) = \text{Cov}(h) = \frac{\sigma^2}{2^{\nu - 1} \cdot \Gamma(\nu)} \cdot \left(\frac{h}{\beta}\right)^{\nu} \cdot K_{\nu}\left(\frac{h}{\beta}\right),
$$
such that $\Gamma(\cdot)$ is the gamma function, and $K_{\nu}(\cdot)$ is the modified Bessel function of the second order.

### Example

For instance, we can compute a covariance matrix for the Matérn model with $\sigma^2$, $\nu = 5$, and $\beta = 5$ in the following way
```{r}
C  <- compute_cov_matrix(points = z, cov_func = "matern_model", sig2 = 1, nu = 5, beta = 5)
print(C[1:5, 1:5])
```

And we can simulate a (zero-mean) Gaussian process with covariance matrix `C` using the `mvrnorm()` function (from the `MASS` package) as follows
```{r}
library(MASS)
GP <- MASS::mvrnorm(n = 1, mu = rep(0, nrow(C)), Sigma = as.matrix(C))
GP <- matrix(data = GP, nrow = length(y), ncol = length(x), byrow = TRUE)

library(pheatmap)
pheatmap::pheatmap(GP, cluster_rows = FALSE, cluster_cols = FALSE)
```


